\bf Fast, Scalable Phrase-Based SMT Decoding

### 2016-09-10 amta.2016/amta2016.tex
Minor fixes, mostly typos and grammatical errors

For many users, decoding is the most time-critical part of the translation process. Making use of the multiple cores that are now ubiquitous in todays servers is a common strategy to ameliorate this issue. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}. That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores. There have been speculation on the causes of the inefficiency as well as potential remedies.

@@

~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describes fast, efficient datastructures for language models. ~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\cite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.

@@

This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers. Unlike ~\cite{mfernandez2016boosting}, we will optimize decoding speed by looking inside the black box. We will compare multicore performance the best-of-breed phrase-table described in ~\cite{junczys_tsd_2012b} with our own implementation. We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available and a framework exists to accommodate other decoding algorithms in future. We use KenLM~\citep{Heafield-kenlm} due to it's popularity and consistent performance, but as with Moses, other language model implementations can be added later.

@@

In the phrase-based model, each rule translates a contiguous sequence of source words. Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit. The target output is constructed strictly left-to-right from the target side from the series of translation rules.

@@

Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event. This also increase speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.

@@

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. However, the model remain the same under the log-linear framework, including having its own weights. %The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.

@@

Over 24\% of the Moses decoder running time is spent on memory management, Table~\ref{tab:optimizing-memory}. This increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} present the tradeoff between decoding time and average model, created by varying the cube-pruning pop-limit. None of the different stack configurations significantly outperform the others in either quality or decoding speed. However, the coverage \& end position produces slightly higher model scores at higher pop-limits, therefore, we continue to use this configuration throughout the rest of this paper.

@@

Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{more-cores.png}} 
\end{tabular}
\caption{Decoding speed with more cores}
\label{fig:more-cores}
\end{figure}

@@

We have presented a new decoder that is compatible with Moses. By studying the shortcomings of the current implementation, we are able to optimize for speed, particularly for multicore operation. This resulted in double digit gains compared to Moses on the same hardware. Our implementation is also unaffected by scalability issues that has afflicted Moses. % and continue to scale well for all possible cores we have tested it on.

>>>>

For many users, decoding is the most time-critical part of the translation process. Making use of the multiple cores that are now ubiquitous in today's servers is a common strategy to ameliorate this issue. However, it has been noticed that the Moses decoder, amongst others, is unable to efficiently use multiple cores~\citep{mfernandez2016boosting}. That is, decoding speed does not substantially increase when more cores are used, in fact, it may actually \emph{decrease} when using more cores. There has been speculation on the causes of the inefficiency as well as potential remedies.

@@

~\cite{Heafield-kenlm} and ~\cite{yasuhara-EtAl:2013:EMNLP} describe fast, efficient data structures for language models. ~\cite{zens2007efficient} describes an implementation of a phrase-table for an SMT decoder that is loaded on demand, reducing the initial loading time and memory requirements. ~\cite{junczys_tsd_2012b} extends this by compressing the on-disk phrase table and lexicalized re-ordering model.

@@

This paper not only focuses on faster single-threaded decoding but also on overcoming the shortcomings of existing decoding implementations on multicore servers. Unlike ~\cite{mfernandez2016boosting}, we will optimize decoding speed by looking inside the black box. We will compare multicore performance the best-of-breed phrase-table described in ~\cite{junczys_tsd_2012b} with our own implementation. We will use the cube-pruning algorithm, however, the standard phrase-based decoding algorithm is also available and a framework exists to accommodate other decoding algorithms in future. We use KenLM~\citep{Heafield-kenlm} due to its popularity and consistent performance, but as with Moses, other language model implementations can be added later.

@@

In the phrase-based model, each rule translates a contiguous sequence of source words. Successive applications of translation rules do not have to be adjacent on the source side, depending on the distortion limit. The target output is constructed strictly left-to-right from the target side using this series of translation rules.

@@

Accumulating unused objects in the memory pools can result in unacceptably high memory usage so object queues are available for high-churn objects which allows the decoder to re-cycle unused objects before the reset event. This also increases speed as LIFO queues are used so that the most recently accessed memory are used, increasing CPU cache hits.

@@

However, the model's query keys are the source and target phrase of each translation rule. Rather than storing the lexicalized re-ordering model separately, we shall integrating it into the translation model, eliminating the need to query a separate file. However, the model remains the same under the log-linear framework, including having its own weights. %The search algorithm will produce identical results to the Moses baseline apart from edge case training configurations, for example, where the lexicalized reordering model is trained on a different dataset than the translation model.

@@

Over 24\% of the Moses decoder running time is spent on memory management (Table~\ref{tab:optimizing-memory}). This increases to 39\% when 32 threads are used, dampening the scalability of the decoder. By contrast, our decoder spends 11\% on memory management and does not significantly increase with more threads.
\begin{table}
\begin{center}
\small
\begin{tabular}{|l|c|c|c|c|} \hline
		& \multicolumn{2}{c|} {Moses}	& \multicolumn{2}{c|} {Our Work} \\ \hline
\# threads	& 1 		& 32	& 1 		& 32  \\ \hline
Memory  	& 24\%		& 39\% 	& 11\%		& 13\% \\
LM 		& 12\%	 	& 2\% 	& 47\%		& 38\% \\ 
Phrase-table	& 9\%	 	& 5\% 	& 2\%		& 4\% \\ 
Lex RO 		& 8\%	 	& 2\% 	& 2\%		& 2\% \\ 
Search 		& 2\%	 	& 0\% 	& 14\%		& 19\% \\ 
Misc/Unknown	& 45\%	 	& 39\% 	& 24\%		& 29\% \\ \hline
\end{tabular}
\end{center}
\caption{Profile of \%age decoding time}
\label{tab:optimizing-memory}
\end{table}

@@

%Table~\ref{tab:stack-configuration} and 
Figure~\ref{fig:stack-configuration} presents the tradeoff between decoding time and average model scores, created by varying the cube-pruning pop-limit. None of the different stack configurations significantly outperform the others in either quality or decoding speed. However, the coverage \& end position produces slightly higher model scores at higher pop-limits, therefore, we continue to use this configuration throughout the rest of this paper.

@@

Our decoder show no scalability issues when we tested with the same model and tested set on a larger server, Figure~\ref{fig:more-cores}.
\begin{figure}
\centering
\begin{tabular}{cc}
{\includegraphics[scale=0.4]{more-cores.png}} 
\end{tabular}
\caption{Decoding speed of our decoder with more cores}
\label{fig:more-cores}
\end{figure}

@@

We have presented a new decoder that is compatible with Moses. By studying the shortcomings of the current implementation, we are able to optimize for speed, particularly for multicore operation. This resulted in double digit gains compared to Moses on the same hardware. Our implementation is also unaffected by scalability issues that have afflicted Moses. % and continue to scale well for all possible cores we have tested it on.
